Mimicry in Online Conversations: An Exploratory Study of Linguistic Analysis Techniques

### 2016-04-25 papers/asonam2016/paper.tex
More progress

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Interactive Alignment Model is the prevailing theory of
[I probably need some help phrasing this].
In this study we investigate how several linguistic alignment techniques can be used
to detect mimicry at the lexical, syntactic and situational levels of the
Interactive Alignment Model.

@@

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% no \IEEEPARstart

Mimicry in online conversations is an important area of research [something about insider threats
and maybe something about improving outcomes of tasks, using mimicry in AI, etc.]

There are a number of measures designed to capture mimicry in text, [something something], and some
efforts have been made to compare and trial these techniques. However, few of these techniques are 
grounded what cognitive science knows about how mimicry works within humans.

This perspective is important so that each measure is grounded in reality.
Moreover, it suggests [does it?] that the best approach may be to combine existing measures to create
a more holistic measure. [don't like any of this wording].

Of the measures discussed in this paper, only one has attempted to measure alignment with this in mind
and some, such as LSM seem to work outside of the theory altogether.

@@

[Not sure what else to put here. Section II does this, section III looks at that type thing?]

\section{Related Work}
This paper draws from many sources for its measures. The most similar piece of work is a comparison
of several methods\cite{xu2015evaluation}, some discussed in this paper. It does not, however, try
to compare the accuracy of the measures, rather, it [does some stuff I don't understand that well,
seems like it looks at some statistical properties?]

[Not sure what else to mention here]

@@

\begin{enumerate}
	\item The situation model --- a representation of the situation being discussed
	\item The semantic representation --- the meaning of the utterance
	\item The syntactic representation --- the sentence structure
	\item The lexical representation --- the choice of words used
	\item The phonological representation --- [I have no idea how to describe the difference between
		phonology and phonetics]
	\item The phonetic representation --- something
\end{enumerate}

The phonological and phonetic layers are ignored by this paper as they do not relate to written text.

The theory states that alignment "percolates" between layers such that increased alignment in one layer
will lead to increased alignment in other layers.

@@

\section{Method}

\subsection{Data}
The dataset consists of 103 dyadic conversations taken from Twitter. Tweets were found via Twitter's 
search API, using English language and being a reply as criteria. We then work up the chain of tweets 
until we hit the end, or a third interlocutor is found, at which point the conversation is cut off and 
kept, as long as there are no fewer than 10 tweets in the conversation. 200 tweets were found using 
this method. Of these, 62 were removed as they were not true conversations (either monologues or 
collaborative fiction writing/roleplay). 18 conversations were removed as they had significant amounts 
of non-English text, and 17 conversations involved bots and were also removed.

@@

The data was classified by three PhD students at Lancaster University with backgrounds in mimicry. Each 
conversation was classified for lexical, syntactic and semantic mimicry on a three point scale where 0 
represents no mimicry, 1 represents low levels of mimicry at that layer, and 2 represents high levels 
of mimicry.

Each measure, with exceptions noted in following subsections [Probably just HAM at this point],
is then run against the data such that each message from author A is compared with each message from 
author B that occurred before the original message.
\footnote{[I could probably make my github repo public and put a link to the code here?]}

[Need to justify dataset and discuss limitations]

\subsection{Lexical}
Lexical stuff.

\subsubsection{LILLA}
Lexical Indiscriminate Local Linguistic Alignment (LILLA)\cite{wang2014linguistic, xu2015evaluation}
is a measure that explicitly works at the lexical layer. It gives the probability that a word from the 
prime document will appear at a particular position in the target document. It is defined as:

\subsubsection{HAM}
The Hierarchical Alignment Model (HAM)\cite{doyle2016robust} ...

\subsection{Syntactic}
Syntactic stuff.

\subsubsection{SILLA}
Syntactic Indiscriminate Local Linguistic Alignment (SILLA)

\subsubsection{SHAM}
SHAM

\subsection{Semantic}
Semantic

\subsubsection{word2vec}
word2vec\cite{mikolov2013distributed}

\subsubsection{doc2vec}
doc2vec\cite{le2014distributed}.

\subsection{Other Measures}
Other

\subsubsection{Linguistic Style Matching}
LSM\cite{niederhoffer2002linguistic}

\subsubsection{Zelig}
Zelig\cite{jones2014finding}

@@

\section{Conclusion and Future Work}
The conclusion goes here.

@@

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
The Interactive Alignment Model is the prevailing theory of
[I probably need some help phrasing this].
In this study we investigate how several linguistic alignment techniques can be used
to detect mimicry at the lexical, syntactic and situation levels of the
Interactive Alignment Model.

@@

\IEEEpeerreviewmaketitle

\section{Introduction}

Mimicry in online conversations is an important and active area of research. Humans tend to mimic each
other in many areas, such as speech
\cite{levelt1982surface, bock1986syntactic, bock1989closed, garrod1987saying},
physical actions\cite{bernieri1988coordinated}
and written text{\cite{danescu2011chameleons, scissors2008linguistic}.
People that mimic each other's language have more successful outcomes in joint tasks
\cite{maddux2008chameleons}. Additionally, mimicry can be an indicator of insider threats
\cite{taylor2013detecting}.

There are a number of measures designed to capture mimicry in text, and some efforts have been made to
compare and trial these techniques. However, few of these techniques are grounded what cognitive 
science knows about how mimicry works within humans.

This perspective is important so that each measure is grounded in reality.
Moreover, it suggests that the best approach may be to combine existing measures to create
a more holistic technique for mimicry detection.

The most similar piece of work is a comparison of several methods\cite{xu2015evaluation}, some 
discussed in this paper. However, it does not try to directly compare the accuracy of the measures,
nor does it look at how they fit into the Interactive Alignment Model.

@@

The novel contributions of this paper are:
\begin{itemize}
	\item Compares existing mimicry detection techniques and how they map to the
		  Interactive Alignment Model
	\item Trials techniques that have not been used for mimicry detection before.
	\item Looks at creating hybrid techniques, combining measures together to find if they are more 
		  accurate when combined. 
\end{itemize}

This paper is divided into six sections. The next section introduces the
Interactive Alignment Model, the psychological model of mimicry we are testing our techniques against.
Section III gives a background of each measure tested in the paper, and defines which layer of the
IAM (if any) the measure falls into.
Section IV addresses the data collected and how it is processed for each technique.
We follow in section V with a discussion of the results, and finally give our conclusions and
suggestions for further work in section VI.

@@

\begin{enumerate}
	\item The situation model --- a representation of the situation being discussed
	\item The semantic representation --- the meaning of the utterance
	\item The syntactic representation --- the sentence structure
	\item The lexical representation --- the choice of words used
	\item The phonological representation --- ???
	\item The phonetic representation --- the sounds of an utterance
\end{enumerate}

The phonological and phonetic representations are not discussed in this paper as they relate purely
to the sound of speech, representations that do not exist in the written text of online conversations.

Further, the situation model is not included as there are currently no measures that can be used for
finding alignment at the situational level.

The theory states that alignment is automatic, that conversation partners will align their speech 
towards each other at every level of the model. In addition, alignment "percolates" between layers such
that increased alignment in one layer will lead to increased alignment in other layers.

@@

\section{Analysis Techniques}

\subsection{Local Linguistic Alignment}
Local Linguistic Alignment (LLA) is a simple,  dual purpose method that can work at both the lexical
and syntactic layers\cite{wang2014linguistic, xu2015evaluation}.

Lexical Indiscriminate Local Linguistic Alignment (LILLA) works at the lexical layer. It gives the
probability that a word from the prime document will appear at a particular position in the target 
document. It is defined as:

Syntactic Indiscriminate Local Linguistic Alignemnt (SILLA) works at the syntactic layer. It works
in exactly the same way as LILLA, but rather than compare words, each sentence is annotated with a
phrase structure tree, and LLA is run on these instead. As such, it measures the probability of a
phrase structure subtree in the prime document appearing occurring at a given position in the target
document.

\subsection{Hierarchical Alignment Model}
\cite{doyle2016robust}

\subsection{Word Vectors}
word2vec\cite{mikolov2013distributed}
doc2vec\cite{le2014distributed}.

\subsection{Linguistic Style Matching}
\cite{niederhoffer2002linguistic}

\section{Method}
The dataset consists of 103 dyadic conversations taken from Twitter. Tweets were found via Twitter's 
search API, using English language and being a reply as criteria. We then work up the chain of tweets 
until we hit the end, or a third interlocutor is found, at which point the conversation is cut off and 
kept, as long as there are no fewer than 10 tweets in the conversation. 200 tweets were found using 
this method. Of these, 62 were removed as they were not true conversations (either monologues or 
collaborative fiction writing/roleplay). 18 conversations were removed as they had significant amounts 
of non-English text, and 17 conversations involved bots and were also removed.

@@

[LIMITATIONS]

The data was classified by three researchers with backgrounds in mimicry independent of the authors of 
this paper.
Each  conversation was classified for lexical, syntactic and semantic mimicry on a three point scale 
where 0 represents no mimicry, 1 represents low levels of mimicry at that layer, and 2 represents high 
levels of mimicry. Classifications from each author were averaged together to give each conversation
three scores between 0 and 2.

Each measure is then run against the data such that each message from author A is compared with each 
message from author B that occurred before the original message. The exception for this is HAM,
where we have used the original implementation [need to actually do this and explain it further].
\footnote{[I could probably make my github repo public and put a link to the code here?]}

@@

@@

\section{Conclusion and Future Work}
The conclusion goes here.
Zelig\cite{jones2014finding}
